---
title: "hdrapi in action"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hdrapi in action}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Easily access the Human Develpment Report Data via API

API documentation: https://hdr.undp.org/data-center/documentation-and-downloads

API GitHub: [HDRData](https://github.com/HDRData/api)

```{r setup}
library(hdrapi)
```


## Get your API key

Described in the following [manual](https://hdr.undp.org/sites/default/files/2023-24_HDR/HDRO_data_api_manual.pdf)

Store your api key in your .Rprofile as described [here](https://stackoverflow.com/questions/37700436/how-to-store-api-key-in-rprofile)

## Get an overview of the API

Get a list of all indicators and search for a keyword:

```{r metadata_1}

indicator_list = get_meta(value = 'Indicators',apikey =HDR_API)

indicator_list[grep('education', indicator_list$name, ignore.case = T),]

```

Get a list the countries and their isocodes. Search the code for a particular country (must be spelled right)

```{r metadata_2}

country_list = get_meta(value = 'Countries', apikey = HDR_API)

country_list[grep('mauritius', country_list$name, ignore.case = T),]

```


More information can be extracted by using the following values `HDGroups`, `HDRegions`, `Dimensions`, `Indices`


## Query examples


```{r get_hdr_data_co2}

get_hdr_data(country_or_aggregation = 'all_countries',
             apikey = HDR_API,
             indicator = 'co2_prod',
             query = 'detailed')

```


Extract Adolescent Birth Rate and CO2 emissions for years 2014 and 2020 for Mauritius, Angola, and the United Arab Emirates

```{r}

get_hdr_data(country_or_aggregation = c('MUS', 'AGO', 'ARE'),
             apikey = HDR_API,
             year = c(2014, 2020),
             indicator = c('abr', 'co2_prod'))
```

Extract all indicators for Mauritius (iso3: MUS) in year 2014

```{r}

get_hdr_data(country_or_aggregation = 'MUS',
             apikey = HDR_API,
             year = 2014)
```


Since no data is available for the year 1950, the following query returns an empty list

```{r get_hdr_data_empty}

get_hdr_data(country_or_aggregation = 'MUS',
             apikey = HDR_API,
             year = 1950)

```

